\documentclass[12pt]{article}
%% Insert Commands in Preface.tex
\usepackage{sagetex}
\input{../Preface.tex}

\begin{document}

\thispagestyle{empty}

{\scshape Huang Linhang; Syx Pek; Qi Ji} \hfill {\scshape \large Differential Geometry} \hfill {\scshape Homework \#5}

\smallskip
\hrule
\bigskip
\section{Section 4.2}

\subsection*{Question 1}


\begin{sagesilent}
    a,b,c,u,v = var('a,b,c,u,v')
    c1 = u*sin(a)*cos(v)
    c2 = u*sin(a)*sin(v)
    c3 = u*cos(a)
\end{sagesilent}

We have that
\begin{enumerate}
\item $\frac{ \partial x}{\partial u} = (\sage{diff(c1,u)}, \sage{diff(c2, u)}, \sage{diff(c3,u)})$ 
\item $\frac{ \partial x}{\partial v} = (\sage{diff(c1,v)}, \sage{diff(c2, v)}, \sage{diff(c3,v)})$
\item $E = \sage{(diff(c1,u)^2 + diff(c2,u)^2 + diff(c3,u)^2).full_simplify()}$
\item $F = \sage{(diff(c1,u)*diff(c1,v) + diff(c2,u)*diff(c2,v) + diff(c3,u)*diff(c3,v)).full_simplify()}$ 
\item $G = \sage{(diff(c1,v)^2 + diff(c2,v)^2 + diff(c3,v)^2).full_simplify()}$
\end{enumerate}

As $E \neq G$, it is not conformal and hence not a local isometry.

To see that it is a local diffeomorphism onto a cone with angle $2\alpha$, observe that for any point $(u, v) \in \mathbb R^2$, construct the set $S = \{(x, y)| x > 0, v-\frac\pi2 \leq y < \frac\pi2\}$. Then $(u, v) \in S \subseteq \mathbb R^2$, and we had that $F|_S$ is a diffeomorphism from $S$ to the cone needed (we had seen this noting that $(u,v)$ is in polar coordinates where $u$ is the radius and $v$ the angle).


\subsection*{Question 4}

Let us consider the stereographic projection $\pi: \mathbb R^2 \to S^2 \setminus{\textbf{N}} \subseteq \mathbb R^3$ where $\textbf{N}$ denotes the north pole.
$$\pi(u, v) = \left(\frac{2u}{1 + u^2 + v^2}, \frac{2v}{1 + u^2 + v^2}, \frac{-1 + u^2 + v^2}{1 + u^2 + v^2} \right).$$

Note that this is sufficient to show local conformality as we can rotate the sphere.
\begin{sagesilent}
    c1 = 2*u/(1 + u^2 + v^2)
    c2 = 2*v/(1 + u^2 + v^2)
    c3 = (-1 + u^2 + v^2)/(1 + u^2 + v^2)
\end{sagesilent}
We had from previous homework that the local chart is a diffeomorphism and 
\begin{enumerate}
\item $\frac{ \partial x}{\partial u} = (\sage{diff(c1,u)}, \sage{diff(c2, u)}, \sage{diff(c3,u)})$ 
\item $\frac{ \partial x}{\partial v} = (\sage{diff(c1,v)}, \sage{diff(c2, v)}, \sage{diff(c3,v)})$
\item $E = \sage{(diff(c1,u)^2 + diff(c2,u)^2 + diff(c3,u)^2).full_simplify()}$
\item $F = \sage{(diff(c1,u)*diff(c1,v) + diff(c2,u)*diff(c2,v) + diff(c3,u)*diff(c3,v)).full_simplify()}$ 
\item $G = \sage{(diff(c1,v)^2 + diff(c2,v)^2 + diff(c3,v)^2).full_simplify()}$
\end{enumerate}

As $E = G$ and $F = 0$, it is a conformal chart.

\subsection*{Question 7}

% so many paired crap in here
\newcommand*{\inner}[1]{\left\langle #1 \right\rangle}
\newcommand*{\set}[1]{\left\{ #1 \right\}}
\newcommand*{\abs}[1]{\left\lvert #1 \right\rvert}

Let \(V\) and \(W\) be (finite-dimensional) vector spaces with inner product denoted by \(\langle,\rangle\)
and let \(F:V\to W\) be a linear map.
Prove that the following conditions are equivalent:
\begin{enumerate}
    \item[a.] \(\inner{F(v_1),F(v_2) } = \inner{v_1,v_2}\) for all \(v_1,v_2\in V\).
    \item[b.] \(\abs{ F(v) } = \abs{ v}\) for all \(v\in V\).
    \item[c.] If \(\set{v_1,\dots,v_n}\) is an orthonormal basis in \(V\), then \(\set{F(v_1),\dots,F(v_n)}\) is an orthonormal basis in \(W\).
    \item[d.] There exists an orthonormal basis \(\set{v_1,\dots,v_n}\) in \(V\) such that \(\set{F(v_1),\dots,F(v_n)}\) is an orthonormal basis in \(W\).
\end{enumerate}

\textbf{(a)$\Rightarrow$(b).} is obvious.

\textbf{(a)$\Rightarrow$(c).} Assume (a), if \(\set{v_1,\dots,v_n}\) is an orthonormal basis in \(V\) then whenever \(i\ne j\),
\(0 = \inner{v_i,v_j} = \inner{F(v_i),F(v_j)}\) and furthermore for each \(i\), by (b) \(\abs{F(v_i)} = \abs{v_i}\).
This shows \(\set{F(v_1),\dots,F(v_n)}\) is orthogonal set, for it to be a basis we need to assume in addition that \(V\) and \(W\) have the same dimension (so \(F\) is surjective).

\textbf{(c)$\Rightarrow$(d).} is trivial as orthonormal bases can always be produced by Gram-Schmidt.

\textbf{(d)$\Rightarrow$(a).} Assume (d), let \(v,v'\in V\).
We can express them in our orthonormal basis as follows
\begin{align*}
    v &= \sum_{i=1}^n c_i v_i,\\
    v' &= \sum_{j=1}^n d_j v_j.
\end{align*}
as \(\set{v_1,\dots,v_n}\) is an orthonormal basis, \(\inner{v_i,v_j} = 0\) whenever \(i\ne j\) and \(\inner{v_i,v_i} = 1\), so
\begin{align*}
    \inner{v,v'} &= \inner{ \sum_{i=1}^n c_i v_i, \sum_{j=1}^n d_j v_j } \\
                 &= \sum_{i=1}^n \sum_{j=1}^n c_i d_j \inner{v_i,v_j} \\
                 &= \sum_{i=1}^n c_i d_i
\end{align*}
and by linearity
\begin{align*}
    F(v) &= \sum_{i=1}^n c_i F(v_i),\\
    F(v') &= \sum_{j=1}^n d_j F(v_j).
\end{align*}
and as \(\set{F(v_1),\dots,F(v_n)}\) is also an orthonormal basis we can perform a similar computation to get
\[ \inner{F(v),F(v')} = \sum_{i=1}^n c_i d_i = \inner{v,v'}  \]
which shows (a).


\subsection*{Question 10}

Let \(S\) be a surface of revolution. Prove that the rotations about its axis are isometries of \(S\).

\emph{Proof.}
Suppose \(S\) is a surface formed by rotating around \(z\)-axis, let \(S\) be parametrised by
\[ \mathbf{x}(u,v) = (\varphi(v)\cos u, \varphi(v)\sin u, \psi(v)) \]
for some \(\varphi, \psi\).
Then the rotation about \(z\)-axis by \(\theta\) can be given by
\[ T = \begin{pmatrix}
    \cos\theta & -\sin\theta & 0 \\
    \sin\theta & \cos\theta & 0 \\
    0 & 0 & 1
\end{pmatrix}. \]
ยง2-3 exercise 11 shows that \(T\) restricted to \(S\) is a diffeomorphism onto \(S\),
we just need to show that it is a local isometry at every point.

Now let \(\overline{\mathbf{x}} = T\circ\mathbf{x}\) and we can compute that
\[ \overline{\mathbf{x}}(u,v) = (\varphi(v)\cos(u+\theta), \varphi(v)\sin(u+\theta), \psi(v)) \]
Note that ยง2-3 Example 4 computes the coefficients of the first fundamental form of \(S\) as
\[ E = \varphi^2,\quad F = 0,\quad G = (\varphi')^2 + (\psi')^2. \]
We can compute \(\overline{E},\overline{F},\overline{G}\) by hand as
\begin{align*}
    \overline{\mathbf{x}}_u &= (-\varphi(v)\sin(u+\theta),\varphi(v)\cos(u+\theta), 0) \\
    \overline{\mathbf{x}}_v &= (\varphi'(v)\cos(u+\theta),\varphi'(v)\sin(u+\theta), \psi'(v)) \\
    \overline{E} &= \left\langle \overline{\mathbf{x}}_u, \overline{\mathbf{x}}_u \right\rangle \\
                 &= \varphi^2 = E\\
    \overline{F} &= \left\langle \overline{\mathbf{x}}_u, \overline{\mathbf{x}}_v \right\rangle \\
                 &= 0 = F \\
    \overline{G} &= \left\langle \overline{\mathbf{x}}_v, \overline{\mathbf{x}}_v \right\rangle \\
                 &= (\varphi')^2 + (\psi')^2 = G
\end{align*}
Applying proposition 1 we have \(\overline{\mathbf{x}}\circ\mathbf{x} = T\) is a local isometry at some arbitrary point, which suffices.

\subsection*{Question 13}
Let $V$ and $W$ be (finite-dimensional) vector spaces with inner products $\langle ,\rangle$. Let $G: V \to W$ be a linear map. Prove that the following conditions are equivalent:
\begin{enumerate}
    \item There exists a real constant $\lambda \neq 0$ such that \begin{equation*}
        \langle G(v_1), G(v_2)\rangle = \lambda^2\langle v_1, v_2\rangle \quad \text{for all} v_1,v_2 \in V.
    \end{equation*}
    \item There exists a real constant $\lambda>0$ such that \begin{equation*}
        |G(v)| = \lambda |v|\quad \text{for all}\quad v \in V.
    \end{equation*}
    \item There exists an orthonormal basis $\{v_1,\dots,v_n\}$ of $V$ such that $\{G(v_1),...,G(v_n)\}$ is an orthogonal basis of $W$ and, also, the vectors $G(v_i), i = 1, ...,n$, have the same (nonzero) length.
\end{enumerate}

If any of these conditions is satisfied, G is called a linear conformal map (or a similitude).\\

\textit{Solution.}
($1 \Rightarrow 2$) We have \begin{align*}
    |G(v)| &= \sqrt{\langle G(v), G(v)\rangle}\\
    &= \sqrt{\lambda^2 \langle v, v\rangle}\\
    &= |\lambda|\sqrt{\langle v, v\rangle}\\
    &= |\lambda||v|,
\end{align*}
where $|\lambda| > 0$ is the positive constant desired.\\

($2 \Rightarrow 1$) We have \begin{align*}
    \langle G(v_1), G(v_2)\rangle &= \frac{1}{2}(|G(v_1)+G(v_2)|^2-|G(v_1)|^2-|G(v_2)|^2),\\
    &= \frac{\lambda^2}{2}(|v_1+v_2|^2-|v_1|^2-|v_2|^2)\\
    &= \lambda^2\langle v_1, v_2\rangle\\
\end{align*}

($1\&2 \Rightarrow 3$) For $\{v_i,v_j\}$ orthonormal we have \begin{align*}
    |G(v_i)| = \lambda|v_i| &= \lambda|v_j| = |G(v_j)|\\
    \langle G(v_i), G(v_j)\rangle &= \lambda^2 \langle v_i, v_j\rangle
\end{align*}
This shows that $\{G(v_1),...,G(v_n)\}$ is an orthogonal basis of $W$ and $G(v_i), i = 1, ...,n$, have the same (nonzero) length.\\

($3 \Rightarrow 2$) For any $v \in V$, let $v = \sum_{i=1}^na_iv_i$. Then \begin{align*}
    |G(v)|^2 &= \langle G(v),G(v)\rangle = \langle \sum_{i=1}^n a_iG(v_i),\sum_{i=1}^n a_iG(v_i)\rangle\\
    &=\sum_{i=1}^n\sum_{i=1}^n a_ia_j\langle G(v_i),G(v_j)\rangle\\
    &=\sum_{i=1}^n a_i^2\langle G(v_i),G(v_i)\rangle + \sum_{i\neq j; i,j\in{1,\dots, n}}a_ia_j\langle G(v_i),G(v_j)\rangle\\
    &=\sum_{i=1}^n \lambda^2 a_i^2\langle v_i, v_j\rangle + 0\\
    &=\lambda^2 \langle v,v\rangle = (|\lambda||v|)^2
\end{align*}
Hence, $|G(v)| = |\lambda||v|$.

\subsection*{Question 16}
Let $\mathbf{x}:U \subset R^2 \to R^3$, where \begin{align*}
    U &=\{(\theta,\varphi)\in R^2:0<\theta<\pi,0<\varphi<2\pi\},\\
    x(\theta,\varphi) &= (\sin{\theta}\cos{\varphi},\sin{\theta}\sin{\varphi},\cos{\theta}),
\end{align*}
be a parametrization of the unit sphere $S^2$. Let\begin{equation*}
     \log\tan \frac{1}{2}\theta = u,\qquad \varphi = v
\end{equation*}
and show that a new parametrization of the coordinate neighborhood $\mathbf{x}(U)= V$ can be given by
\begin{equation*}
    \mathbf{y}(u,v) = (\text{sech } u\cos v,\text{sech } u\sin v,\tanh u).
\end{equation*}
Prove that in the parametrization $y$ the coefficients of the first fundamental form are \begin{equation*}
    E=G=\text{sech}^2~u,\qquad F=0.
\end{equation*}
Thus, $\mathbf{y}^{-1} : V \subset S^2 \to R^2$ is a conformal map which takes the meridians and
parallels of $S^2$ into straight lines of the plane. This is called \textit{Mercator's projection}.\\

\textit{Solution.} We have \begin{equation*}
    \theta = 2\arctan e^u, \qquad v = \varphi.
\end{equation*} Hence,\begin{align*}
    \mathbf{y}(u,v) &= \mathbf{x}(2\arctan e^u,v)\\
    &=(\sin{2\arctan e^u}\cos{v},\sin{2\arctan e^u}\sin{v},\cos{2\arctan e^u})\\
    &=(\text{sech }u\cos{v}, \text{sech }u\sin{v},\tanh u).
\end{align*}
Therefore, we have \begin{align*}
    \mathbf{y}_u &= \langle -\tanh u \text{ sech }u\cos v, -\tanh u \text{ sech }u\sin v, 1 - \tanh^2 u \rangle\\
    \mathbf{y}_v &= (-\text{sech }u\sin{v}, \text{sech }u\cos{v},0)\\
    E &= |\mathbf{y}_u|^2 = \tanh^2 u \text{ sech}^2~u + (1-\tanh^2 u)^2\\
    &= \tanh^2 u (1-\tanh^2 u) + (1-\tanh^2 u)^2\\
    &= 1-\tanh^2 u = \text{sech}^2~u\\
    G &= |\mathbf{y}_v|^2 = \text{sech}^2~u (\sin^2 u+\cos^2 u) = \text{sech}^2~u\\
    F &= \tanh u \text{sech}^2~u\sin v\cos v - \tanh u \text{sech}^2~u\sin v\cos v = 0
\end{align*}


\end{document}
